---
title: "LITMUS Processing — SUSHI Report"
output:
  html_document:
    self_contained: true
    toc: true
    toc_depth: 2
    theme: flatly
    df_print: paged
params:
  params_json: ""
  feature_table: ""
  feature_meta: ""
  sample_meta: ""
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,
                      fig.align = "center", dpi = 100)
suppressPackageStartupMessages({
  library(jsonlite)
  library(knitr)
})

`%||%` <- function(a, b) if (!is.null(a) && length(a) > 0 && nzchar(as.character(a[1]))) a else b

params_info   <- tryCatch(fromJSON(params$params_json), error = function(e) list())
feature_table <- tryCatch(read.csv(params$feature_table, check.names = FALSE,
                                    stringsAsFactors = FALSE),
                           error = function(e) NULL)
feature_meta  <- tryCatch(read.csv(params$feature_meta, stringsAsFactors = FALSE),
                           error = function(e) NULL)
sample_meta   <- tryCatch(read.csv(params$sample_meta,  stringsAsFactors = FALSE),
                           error = function(e) NULL)

# Separate the leading Sample column from the intensity matrix.
if (!is.null(feature_table) && ncol(feature_table) >= 2) {
  sample_col   <- feature_table[[1]]
  intensity_df <- feature_table[, -1, drop = FALSE]
  intensity_m  <- as.matrix(intensity_df)
  storage.mode(intensity_m) <- "numeric"
  rownames(intensity_m) <- sample_col
} else {
  intensity_m <- NULL
}
```

## Run summary

```{r summary}
if (length(params_info) > 0) {
  s <- data.frame(
    Field = c("Tool version", "Timestamp (UTC)", "Engine", "Demo mode",
              "Polarity filter", "Sample types",
              "# samples", "# features", "Groups", "Annotation",
              "LITMUS path"),
    Value = c(
      params_info$version              %||% "",
      params_info$timestamp            %||% "",
      params_info$engine               %||% "",
      as.character(params_info$demo_mode %||% ""),
      params_info$filters$polarity     %||% "",
      paste(params_info$filters$sample_types %||% "", collapse = ", "),
      params_info$n_samples            %||% "",
      params_info$n_features           %||% "",
      paste(params_info$groups         %||% "", collapse = ", "),
      params_info$annotation           %||% "",
      params_info$litmus_path          %||% ""
    ),
    stringsAsFactors = FALSE
  )
  kable(s, align = "ll")
} else {
  cat("(parameters.json unavailable)")
}
```

## Engine parameters

```{r engine-params}
if (!is.null(params_info$engine_params)) {
  ep <- params_info$engine_params
  df <- data.frame(
    Parameter = names(ep),
    Value     = vapply(ep, function(x) paste(x, collapse = ", "), character(1)),
    stringsAsFactors = FALSE,
    row.names = NULL
  )
  kable(df, align = "ll")
}
```

## Sample metadata

```{r sample-meta}
if (!is.null(sample_meta)) {
  kable(sample_meta, align = "l", caption = "Sample metadata (from dataset.tsv)")
}
```

## Per-group sample count

```{r group-counts}
if (!is.null(sample_meta)) {
  tab <- as.data.frame(table(
    Group      = sample_meta$group,
    SampleType = sample_meta$sample_type,
    Polarity   = sample_meta$polarity
  ))
  tab <- tab[tab$Freq > 0, ]
  kable(tab, row.names = FALSE, caption = "Sample distribution (Grouping × Sample Type × Polarity)")
}
```

## Feature table preview

```{r feature-preview}
if (!is.null(intensity_m)) {
  n_show_feat <- min(10, ncol(intensity_m))
  n_show_samp <- min(10, nrow(intensity_m))
  preview <- intensity_m[seq_len(n_show_samp), seq_len(n_show_feat), drop = FALSE]
  cat("Matrix dimensions: **", nrow(intensity_m), " samples × ",
      ncol(intensity_m), " features**\n\n", sep = "")
  kable(round(preview, 2),
        caption = sprintf("First %d samples × %d features", n_show_samp, n_show_feat))
}
```

## Feature metadata (first 20)

```{r feature-meta}
if (!is.null(feature_meta)) {
  kable(head(feature_meta, 20), caption = "First 20 features")
}
```

## Intensity distribution

```{r dist-plot, fig.width=8, fig.height=4}
if (!is.null(intensity_m) && nrow(intensity_m) > 0 && ncol(intensity_m) > 0) {
  log_vals <- log10(as.vector(intensity_m[intensity_m > 0]))
  par(mar = c(4, 4, 2, 1))
  hist(log_vals, breaks = 40,
       main = "log10(intensity) across all samples × features",
       xlab = "log10(intensity)", col = "#74a9cf", border = "white")
}
```

## PCA on log-scaled matrix

```{r pca, fig.width=7, fig.height=5}
if (!is.null(intensity_m) && nrow(intensity_m) >= 3 && ncol(intensity_m) >= 3) {
  # log10 + small offset, then PCA on samples (rows)
  m <- log10(intensity_m + 1)
  keep_feat <- apply(m, 2, sd, na.rm = TRUE) > 1e-10
  if (sum(keep_feat) >= 2) {
    m <- m[, keep_feat, drop = FALSE]
    pca <- prcomp(m, center = TRUE, scale. = FALSE)
    ve  <- (pca$sdev ^ 2) / sum(pca$sdev ^ 2) * 100
    grp <- if (!is.null(sample_meta)) sample_meta$group[match(rownames(m),
                                                              sample_meta$sample_id)]
           else rep("all", nrow(m))
    grp[is.na(grp)] <- "unknown"
    cols <- rainbow(length(unique(grp)))[as.integer(factor(grp))]
    par(mar = c(4, 4, 2, 8), xpd = TRUE)
    plot(pca$x[, 1], pca$x[, 2],
         col = cols, pch = 19, cex = 1.5,
         xlab = sprintf("PC1 (%.1f%%)", ve[1]),
         ylab = sprintf("PC2 (%.1f%%)", ve[2]),
         main = "PCA — log10 intensities, mean-centered")
    legend("topright", inset = c(-0.25, 0),
           legend = unique(grp), col = rainbow(length(unique(grp))),
           pch = 19, bty = "n", cex = 0.8)
  } else {
    cat("Not enough variable features for PCA.\n")
  }
}
```

## Next step

Feed `feature_table.csv` + `sample_meta.csv` into `LitmusNormalizeApp` (or a
future LitmusStatsApp) to continue the LITMUS 13-step pipeline. The
`sample_meta.csv` already carries the `group` column as metadata, so the
downstream app will NOT need to parse filenames.

---
*Generated by `litmus_process.R` via `LitmusProcessApp.rb` (branch
`litmus-sushi-app`, PoC 2026-04-24). Backend: LITMUS
`scripts/R/processing_engine.R`.*
