---
title: "LITMUS Lipidomics Normalization — SUSHI Report"
output:
  html_document:
    self_contained: true
    toc: true
    toc_depth: 2
    number_sections: false
    theme: flatly
params:
  params_json: ""
  norm_csv:    ""
  qc_feats:    ""
  qc_samples:  ""
  final_scale: "raw"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE,
                      fig.width = 9, fig.height = 5)
library(jsonlite)
library(ggplot2)

# base-R fallback for rlang's null-default operator (Rmd may not autoload rlang)
`%||%` <- function(a, b) if (!is.null(a) && length(a) > 0) a else b

cfg <- if (nzchar(params$params_json) && file.exists(params$params_json)) {
  fromJSON(params$params_json)
} else list()

norm_mat <- if (nzchar(params$norm_csv) && file.exists(params$norm_csv)) {
  read.csv(params$norm_csv, check.names = FALSE)
} else data.frame()

qc_feat <- if (nzchar(params$qc_feats) && file.exists(params$qc_feats)) {
  read.csv(params$qc_feats, check.names = FALSE)
} else data.frame()

qc_samp <- if (nzchar(params$qc_samples) && file.exists(params$qc_samples)) {
  read.csv(params$qc_samples, check.names = FALSE)
} else data.frame()

final_scale <- params$final_scale
```

# Summary

```{r}
cat("**Run:** ", format(Sys.time(), "%Y-%m-%d %H:%M:%S"), "\n\n", sep = "")
cat("**Samples:** ", cfg$n_samples %||% NA, "\n\n", sep = "")
cat("**Features:** ", cfg$n_features %||% NA, "\n\n", sep = "")
cat("**Normalization:** ", cfg$method %||% NA, " → **Transform:** ",
    cfg$transform %||% NA, " → **Scaling:** ", cfg$scaling %||% NA, "\n\n", sep = "")
cat("**Final data scale:** ", final_scale, "\n\n", sep = "")
```

# Pipeline Parameters

```{r}
if (length(cfg$pipeline_steps) > 0) {
  knitr::kable(as.data.frame(do.call(rbind, lapply(cfg$pipeline_steps, as.data.frame))),
               caption = "Pipeline steps applied (in order)")
} else {
  cat("(No pipeline steps — identity pass.)")
}
```

# Per-Sample Statistics

```{r}
if (nrow(qc_samp) > 0) {
  knitr::kable(qc_samp, digits = 3, caption = "Per-sample statistics (mean, SD by group)")
} else {
  cat("No sample-level stats available.")
}
```

```{r fig.cap="Sample mean by group."}
if (nrow(qc_samp) > 0) {
  ggplot(qc_samp, aes(x = sample, y = mean, fill = group)) +
    geom_col() +
    coord_flip() +
    theme_minimal(base_size = 11) +
    labs(x = NULL, y = paste0("Mean intensity (", final_scale, "-scale)"),
         title = "Per-sample mean")
}
```

# Per-Feature Statistics

```{r}
if (nrow(qc_feat) > 0) {
  knitr::kable(head(qc_feat, 15), digits = 3,
               caption = "Top 15 features (alphabetical). Full table: qc_metrics_features.csv")
}
```

```{r fig.cap="Feature-level dispersion distribution."}
if (nrow(qc_feat) > 0) {
  metric <- if (final_scale == "raw" && "cv_pct" %in% names(qc_feat)) "cv_pct" else "sd"
  ggplot(qc_feat, aes(x = .data[[metric]])) +
    geom_histogram(bins = 30, fill = "#003c68", color = "white") +
    theme_minimal(base_size = 11) +
    labs(x = if (metric == "cv_pct") "CV%" else paste0("SD (", final_scale, "-scale)"),
         y = "Feature count",
         title = paste0("Distribution of per-feature ",
                        if (metric == "cv_pct") "CV%" else "SD"))
}
```

# Output Files

- `normalized_matrix.csv` — samples × features matrix after the full pipeline
- `qc_metrics.csv` / `qc_metrics_features.csv` — per-feature stats
- `qc_metrics_samples.csv` — per-sample stats
- `parameters.json` — parameters used (for provenance)
- `00index.html` — this report

# Provenance

```{r}
if (length(cfg) > 0) {
  cat("**Input:** `", cfg$input %||% "", "`\n\n", sep = "")
  cat("**LITMUS path:** `", cfg$litmus_path %||% "", "`\n\n", sep = "")
  cat("**Timestamp:** `", cfg$timestamp %||% "", "`\n\n", sep = "")
  cat("**Tool version:** `", cfg$version %||% "", "`\n\n", sep = "")
}
```

---

<small>Generated by **LitmusNormalizeApp** (SUSHI PoC) — FGCZ</small>
